---
title: "Proposal title"
subtitle: "Proposal"
format: html
output:
   warning: false
   message: false
   error: false
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
### GETTING THE LIBRARIES
devtools::install_github("choonghyunryu/dlookr")

if (!require(pacman))
  install.packages(pacman)

pacman::p_load(tidyverse,
               tidytuesdayR,
               dlookr,
               formattable)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

bigtech <- tidytuesdayR::tt_load('2023-02-07')
big_tech_stock_prices <- bigtech$big_tech_stock_prices
big_tech_companies <- bigtech$big_tech_companies

```

The companies that are study for our data analysis are Apple, Adobe, Amazon.com, Salesforce, Cisco Systems, Alphabet, IBM, Intel, Meta, Microsoft, Netflix, NVIDIA, Oracle and TESLA.

### Stock Prices Dataset

```{r}
#| label: load-dataset1
#| message: false
big_tech_stock_prices |>
  diagnose() |>
  formattable()
```

The dataset contains 14 big tech companies daily stock prices with a total of approx. 45088 observations. There is no missing data in the observation which helps to use all the columns available in the data set.

[***Variable description***]{.underline}

| variable     | class     | description                                                                                                                                                                                                                       |
|-------------|-------------|-----------------------------------------------|
| stock_symbol | character | stock_symbol                                                                                                                                                                                                                      |
| date         | double    | date                                                                                                                                                                                                                              |
| open         | double    | The price at market open.                                                                                                                                                                                                         |
| high         | double    | The highest price for that day.                                                                                                                                                                                                   |
| low          | double    | The lowest price for that day.                                                                                                                                                                                                    |
| close        | double    | The price at market close, adjusted for splits.                                                                                                                                                                                   |
| adj_close    | double    | The closing price after adjustments for all applicable splits and dividend distributions. Data is adjusted using appropriate split and dividend multipliers, adhering to Center for Research in Security Prices (CRSP) standards. |
| volume       | double    | The number of shares traded on that day.                                                                                                                                                                                          |

```{r}
#| label: load-dataset2
#| message: false
big_tech_stock_prices |> 
  group_by(stock_symbol) |> 
  summarize(min_date = min(date), max_date = max(date)) |> 
  formattable()
```

The data is available for 12 years except for Meta which starts from year 2012.

### Big Tech Companies Dataset

```{r}
#| label: load-dataset3
#| message: false
big_tech_companies |>
  formattable()
```

Big Tech Companies dataset provides the company names against the stock symbols.

### Why did we choose this data set?

While stock prices are a familiar dataset, our selection is driven by a careful consideration of both technical and analytical factors. Stock prices, being a culmination of market sentiment and company performance, provide valuable insights. Analyzing its trends and drivers enables investors to make well-informed decisions.

The dataset spans nearly a decade, allowing us to observe trends from the post-financial crisis period to the third wave of the COVID-19 pandemic. By creating impactful visualizations for the top 14 big tech companies, we aim to gain a comprehensive understanding of how the tech industry has been perceived in the market over this time frame.

In essence, our connection to this dataset extends beyond technical considerations; it also delves into analytical aspects, empowering us to make informed decisions within the provided data.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
